{\bf {\Large Title....}}

### 2016-07-11 ident_manu.Rnw
results text done

Response of \ac{do} to parameter perturbations was not uniform acros the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January), the spring/summer increase represents production associated with expected seasonal maxima, and the remaining time series from August to the end of year shows complete removal of \ac{do} from the system as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \Sexpr{sensrngcat$Zooplankton$Parameter[2]}, which caused an increase in \ac{do}.  The phytoplankton parameter, \Sexpr{sensrngcat$Phytoplankton$Parameter[2]}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\Sexpr{sensrngcat$Optics$Parameter[2]}) and phytoplankton (\Sexpr{sensrngcat$Phytoplankton$Parameter[2]}) parameters.

@@

<<identres>>=
data(cat_ident)
data(top_ident)

@@

<<echo = F>>=
temp_coll <- par_txt(c('Tref(nospA+nospZ)_1', 'Tref(nospA+nospZ)_2', 'Tref(nospA+nospZ)_3', 'Tref(nospA+nospZ)_4', 'Tref(nospA+nospZ)_5', 'Tref(nospA+nospZ)_6', 'Tref(nospA+nospZ)_7', 'Tref(nospA+nospZ)_8'))
@

@@

Comparison of identifiability between categories showed that phytoplankton and zooplankton had the least identifiable parameter subsets.  As noted above, FishTank includes six phytoplankton and two zooplankton groups to characterize community structure and foodweb dynamics that likely have an important role in hypoxia development.  However, structural equations for each group do not vary considerably such that variation in parameter values primarily control differences between the groups, e.g., large-bodied vs small-bodied plankton, slow-growing vs. fast-growing plankton.  To obtain identifiability estimates of the plankton categories that were independent of groups, the identifiability analyses were re-evaluated using only one phytoplankton and one zooplankton group (\cref{fig:p1z1identbox}).

Reading each plot from left to right can be interpreted as including additional parameters, where each parameter is ranked by relative sensitivity.  The inset in each plot shows the identifiability of including parameters, up to a maximum where additional inclusion exceeds the identifability threshold of fifteen.  The scenarios for including parameters all begin with the parameters that have the greatest effect on the model output.  The inclusion of additional parameters depends on the scenario.  The first scenario selected parameters by decreasing sensitivity within each category (i.e., four separate models calibrated for optics, organics, phytoplankton, or zooplankton), the second scenario selected parameters by sensitivity regardless of category, and the third scenario selects parameters by sensitivity with equal representation between categories.

@@

@@

% parameter exclusion temp
<<excltemp, fig.cap = 'Identifiability (as $\\gamma$, \\cref{gameq}) of temperature parameters for subset combinations in \\cref{fig:identbox}.  Identifiability is evaluated for subsets that excluded and included the parameters at the top of each plot. Identifiability of including all seven parameters is in \\cref{fig:identbox}. Grey lines indicate potential thresholds at $\\gamma = 10, 15$ for maximum acceptable identifiability.', fig.height = 8, fig.width = 5, echo = F, cache = TRUE, out.width = '0.6\\textwidth'>>=
data(cat_ident)

>>>>

Response of \ac{do} to parameter perturbations was not uniform acros the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January), the spring/summer increase represents production associated with expected seasonal maxima, and the remaining time series from August to the end of year shows complete removal of \ac{do} from the system as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the biogeochemical model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \Sexpr{sensrngcat$Zooplankton$Parameter[2]}, which caused an increase in \ac{do}.  The phytoplankton parameter, \Sexpr{sensrngcat$Phytoplankton$Parameter[2]}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\Sexpr{sensrngcat$Optics$Parameter[2]}) and phytoplankton (\Sexpr{sensrngcat$Phytoplankton$Parameter[2]}) parameters.

@@

<<>>=
data(cat_ident)
data(top_ident)

@@

<<>>=
temp_coll <- par_txt(c('Tref(nospA+nospZ)_1', 'Tref(nospA+nospZ)_2', 'Tref(nospA+nospZ)_3', 'Tref(nospA+nospZ)_4', 'Tref(nospA+nospZ)_5', 'Tref(nospA+nospZ)_6', 'Tref(nospA+nospZ)_7', 'Tref(nospA+nospZ)_8'))
@

@@

<<>>=
data(p1z1cat_ident)
data(p1z1top_ident)

# percent combos less than gamma 15, 10 by category
p1z1collcatpers <- lapply(p1z1cat_ident, function(x){

# percent combos less than gamma 15, 10 by top in each category
p1z1colltoppers <- lapply(p1z1top_ident, function(x){

  })
@

Comparison of identifiability between categories showed that phytoplankton and zooplankton had the least identifiable parameter subsets.  As noted above, FishTank includes six phytoplankton and two zooplankton groups to characterize community structure and foodweb dynamics that likely have an important role in hypoxia development.  However, structural equations for each group do not vary considerably such that variation in parameter values primarily control differences between the groups, e.g., large-bodied vs small-bodied plankton, slow-growing vs. fast-growing plankton.  To obtain identifiability estimates of the plankton categories that were independent of groups, the identifiability analyses were re-evaluated using only one phytoplankton and one zooplankton group (\cref{fig:p1z1identbox}). Compared to all groups, evaluating a single group improved identifiability such that a majority of parameter combinations were below the threshold (\Sexpr{p1z1collcatpers$Phytoplankton[1]}\% less than $\gamma = 15$, \Sexpr{p1z1collcatpers$Phytoplankton[2]}\% less than $\gamma = 10$ for Phytoplankton; \Sexpr{p1z1collcatpers$Zooplankton[1]}\%, less than $\gamma = 15$, \Sexpr{p1z1collcatpers$Zooplankton[2]}\%, less than $\gamma = 10$ for zooplankton).  Analysis of identifiability for top parameters in all categories did now show similar improvements in identifiability.  Although all combinations for the top one, two, and three parameters in each category were well below the threshold, selecting the top four and five parameters showed an increase (\Sexpr{p1z1colltoppers$top4[1]}\% less than $\gamma = 15$, \Sexpr{p1z1colltoppers$top4[2]}\% less than $\gamma = 10$) and a decrease (\Sexpr{p1z1colltoppers$top5[1]}\% less than $\gamma = 15$, \Sexpr{p1z1colltoppers$top5[2]}\% less than $\gamma = 10$) in identifiability, respectively, compared to results that included all phytoplankton and zooplankton groups.

\subsection{Parameter selection}

<<>>=
data(tocal)

# n parms and ident for each selection heuristic
h1_id <- tocal$cat %>% 
  split(., .$Category) %>% 
  lapply(function(x){
    x <- x[x$coll < 15, ]
    list(nrow(x), scinot(max(x$coll), digits = 1))
  }) 
h2_id <- filter(tocal$top, coll < 15) %>%
  list(nrow(.), scinot(max(.$coll), digits = 1)) %>% 
  .[c(2, 3)]
h3_id <- filter(tocal$cattop, coll < 15) %>%
  list(nrow(.), scinot(max(.$coll), digits = 1)) %>% 
  .[c(2, 3)]

# which parameter broke selection for h2, h3
h2_brk <- tocal$top
sel <- which(h2_brk$coll < 15)
h2_brk <- h2_brk[c(sel, rev(sel)[1] + 1), ] %>% 
  arrange(-coll) %>% 
  .[1:2, ] %>% 
  list(c(as.character(.$Parameter[1]), .$coll)) %>% 
  .[[2]]
# which parameter broke selection for h2, h3
h3_brk <- tocal$cattop
sel <- which(h3_brk$coll < 15)
h3_brk <- h3_brk[c(sel, rev(sel)[1] + 1), ] %>% 
  arrange(-coll) %>% 
  .[1:2, ] %>% 
  list(c(as.character(.$Parameter[1]), .$coll)) %>% 
  .[[2]]
@

Each of the three selection heuristics (blocked by parameter category, independent of category, all categories equally) for \ac{do} differed in the number of selected parameters and distribution of parameters within each category (if applicable, \cref{fig:heurist}).  For the first selection heuristic, all sensitive parameters from the optics ($n = $ \Sexpr{scinot(h1_id$Optics[[1]], digits = 0)}, $\gamma =$  \Sexpr{h1_id$Optics[[2]]}) and organic matter ($n = $ \Sexpr{scinot(h1_id$`Organic Matter`[[1]], digits = 0)}, $\gamma =$  \Sexpr{h1_id$`Organic Matter`[[2]]}) categories were selected, whereas \Sexpr{toeng(h1_id$Phytoplankton[[1]])[[1]]} were selected for phytoplankton ($\gamma = $ \Sexpr{h1_id$Phytoplankton[[2]]}), \Sexpr{toeng(h1_id$Temperature[[1]])[[1]]} for temperature ($\gamma = $ \Sexpr{h1_id$Temperature[[2]]}), and \Sexpr{toeng(h1_id$Zooplankton[[1]])[[1]]} for zooplankton ($\gamma = $ \Sexpr{h1_id$Zooplankton[[2]]}).  For the second selection heuristic, \Sexpr{toeng(h2_id[[1]])[[1]]} parameters were selected ($\gamma = $ \Sexpr{h2_id[[2]]} with a majority from the phytoplankton category.  For the third heuristic, \Sexpr{toeng(h3_id[[1]])[[1]]} parameters were selected ($\gamma = $ \Sexpr{h3_id[[2]]}) with equal representation between categories. For the second and third selection heuristics, an individual parameter caused a disproportianate increase in $\gamma$ that forced the selection to stop.  Selection independent of category showed that including \Sexpr{par_txt(h2_brk[1])} caused in increase in $\gamma$ from \Sexpr{scinot(h2_brk[3], digits = 1)} to \Sexpr{scinot(h2_brk[2], digits = 1)} and equal selection within cateogies showed that including \Sexpr{par_txt(h3_brk[1])} caused in increase in $\gamma$ from \Sexpr{scinot(h3_brk[3], digits = 1)} to \Sexpr{scinot(h3_brk[2], digits = 1)}.  Finally, parameter selection for other state variables (\ac{chla}, \ac{par}, nitrate, ammonium, particulate organic matter, dissolved organic matter, phosphorus) also showed that a limited number of parameters were identifiable (for brevity, only results using the second set of selection heuristics are shown, \cref{fig:heurist_stts}).  Less than ten parameters were selected for the remaining state variables with most selected from the phytoplankton category.  Interestingly, no parameters from the temperature category were selected.

@@

Why did identifiability decrease for top five parameters in each category after removing redundant phyto/zoop groups?

@@

% parameter exclusion temp
<<excltemp, fig.cap = 'Identifiability (as $\\gamma$, \\cref{gameq}) of temperature parameters for subset combinations in \\cref{fig:identbox}.  Identifiability is evaluated for subsets that excluded and included the parameters at the top of each plot. Identifiability of including all seven parameters is in \\cref{fig:identbox}. Grey lines indicate potential thresholds at $\\gamma = 10, 15$ for maximum acceptable identifiability.', fig.height = 8, fig.width = 5, echo = F, cache = TRUE, out.width = "0.6\\textwidth">>=
data(cat_ident)

### 2016-07-11 ident_manu.tex
results text done

\begin{singlespace}
\title{{\bf {\Large Title....}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^1$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:lehrter.john@epa.gov}{lehrter.john@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Sun Jul 10 15:30:23 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

Response of \ac{do} to parameter perturbations was not uniform acros the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January), the spring/summer increase represents production associated with expected seasonal maxima, and the remaining time series from August to the end of year shows complete removal of \ac{do} from the system as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \textit{ZKa$_{z1}$}, which caused an increase in \ac{do}.  The phytoplankton parameter, \textit{Qc$_{p1}$}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\textit{astar490}) and phytoplankton (\textit{Qc$_{p1}$}) parameters.

@@

Comparison of identifiability between categories showed that phytoplankton and zooplankton had the least identifiable parameter subsets.  As noted above, FishTank includes six phytoplankton and two zooplankton groups to characterize community structure and foodweb dynamics that likely have an important role in hypoxia development.  However, structural equations for each group do not vary considerably such that variation in parameter values primarily control differences between the groups, e.g., large-bodied vs small-bodied plankton, slow-growing vs. fast-growing plankton.  To obtain identifiability estimates of the plankton categories that were independent of groups, the identifiability analyses were re-evaluated using only one phytoplankton and one zooplankton group (\cref{fig:p1z1identbox}).

Reading each plot from left to right can be interpreted as including additional parameters, where each parameter is ranked by relative sensitivity.  The inset in each plot shows the identifiability of including parameters, up to a maximum where additional inclusion exceeds the identifability threshold of fifteen.  The scenarios for including parameters all begin with the parameters that have the greatest effect on the model output.  The inclusion of additional parameters depends on the scenario.  The first scenario selected parameters by decreasing sensitivity within each category (i.e., four separate models calibrated for optics, organics, phytoplankton, or zooplankton), the second scenario selected parameters by sensitivity regardless of category, and the third scenario selects parameters by sensitivity with equal representation between categories.

>>>>

\begin{singlespace}
\title{{\bf {\Large Title....}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^1$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:lehrter.john@epa.gov}{lehrter.john@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Mon Jul 11 09:01:42 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

Response of \ac{do} to parameter perturbations was not uniform acros the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January), the spring/summer increase represents production associated with expected seasonal maxima, and the remaining time series from August to the end of year shows complete removal of \ac{do} from the system as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the biogeochemical model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \textit{ZKa$_{z1}$}, which caused an increase in \ac{do}.  The phytoplankton parameter, \textit{Qc$_{p1}$}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\textit{astar490}) and phytoplankton (\textit{Qc$_{p1}$}) parameters.

@@

Comparison of identifiability between categories showed that phytoplankton and zooplankton had the least identifiable parameter subsets.  As noted above, FishTank includes six phytoplankton and two zooplankton groups to characterize community structure and foodweb dynamics that likely have an important role in hypoxia development.  However, structural equations for each group do not vary considerably such that variation in parameter values primarily control differences between the groups, e.g., large-bodied vs small-bodied plankton, slow-growing vs. fast-growing plankton.  To obtain identifiability estimates of the plankton categories that were independent of groups, the identifiability analyses were re-evaluated using only one phytoplankton and one zooplankton group (\cref{fig:p1z1identbox}). Compared to all groups, evaluating a single group improved identifiability such that a majority of parameter combinations were below the threshold ($56.3$\% less than $\gamma = 15$, $47.1$\% less than $\gamma = 10$ for Phytoplankton; $83.4$\%, less than $\gamma = 15$, $72.6$\%, less than $\gamma = 10$ for zooplankton).  Analysis of identifiability for top parameters in all categories did now show similar improvements in identifiability.  Although all combinations for the top one, two, and three parameters in each category were well below the threshold, selecting the top four and five parameters showed an increase ($84.2$\% less than $\gamma = 15$, $78.7$\% less than $\gamma = 10$) and a decrease ($54.4$\% less than $\gamma = 15$, $51$\% less than $\gamma = 10$) in identifiability, respectively, compared to results that included all phytoplankton and zooplankton groups.

\subsection{Parameter selection}

Each of the three selection heuristics (blocked by parameter category, independent of category, all categories equally) for \ac{do} differed in the number of selected parameters and distribution of parameters within each category (if applicable, \cref{fig:heurist}).  For the first selection heuristic, all sensitive parameters from the optics ($n = $ $4$, $\gamma =$  $3.7$) and organic matter ($n = $ $7$, $\gamma =$  $1.7$) categories were selected, whereas 23 were selected for phytoplankton ($\gamma = $ $13.7$), five for temperature ($\gamma = $ $11.4$), and five for zooplankton ($\gamma = $ $3.1$).  For the second selection heuristic, 20 parameters were selected ($\gamma = $ $4.6$ with a majority from the phytoplankton category.  For the third heuristic, 18 parameters were selected ($\gamma = $ $9.2$) with equal representation between categories. For the second and third selection heuristics, an individual parameter caused a disproportianate increase in $\gamma$ that forced the selection to stop.  Selection independent of category showed that including \textit{KNO3} caused in increase in $\gamma$ from $4.6$ to $15.6$ and equal selection within cateogies showed that including \textit{Tref(nospA+nospZ)$_{p2}$} caused in increase in $\gamma$ from $9.2$ to $66.2$.  Finally, parameter selection for other state variables (\ac{chla}, \ac{par}, nitrate, ammonium, particulate organic matter, dissolved organic matter, phosphorus) also showed that a limited number of parameters were identifiable (for brevity, only results using the second set of selection heuristics are shown, \cref{fig:heurist_stts}).  Less than ten parameters were selected for the remaining state variables with most selected from the phytoplankton category.  Interestingly, no parameters from the temperature category were selected.
